Accelerating adaptive ultrasound algorithms using graphics processing units \vspace{0.5cm}

### 2014-03-22 ThesisJP/Background/background.tex
Finished section on adaptive volume rendering.

The iterative composition scheme for front-to-back ray casting of one ray, from $s_0$ to $s_{N_s-1}$ where $s$ is the sample position and $N_s$ is the number of samples along the ray, is
\begin{align}
I(s_0, s_{i+1}) &= I(s_0, s_{i}) + (1 - \alpha_{\text{acc},i}) c_{i+1}\\
\alpha_{\text{acc},i+1} &= \alpha_{\text{acc},i} + (1 - \alpha_{\text{acc},i}) \alpha(c_i+1).
\label{eq:fronToBack}
\end{align}
Here, $\alpha(v)$ is the opacity transfer function, $c_{k}$ is the color intensity of sample $s_{k}$, $\alpha_{\text{acc}}$ is the accumulated opacity, and $I(s_0, s_{N_s-1})$ is the intensity value of all the projected samples combined. If $\alpha(v)$ is close to one for one or two consecutive samples, $\alpha_{\text{acc}}$ will quickly approach a value of one. An $\alpha_{acc,k}$-value close to one means that the contribution from $c_{k+1}, \cdots, c_{N_s-1}$ to the final value, $I(s_0, s_{N_s-1})$, will be minimal. Hence, the iteration can stop at $i=k$ without loosing visible samples. This optimization technique is known as \textit{early-ray-termination}, and will significantly speed up our rendering algorithm if an opaque material is rendered.

@@

Volume rendering of a single volume has a complexity of $O(N_xN_yN_s)$, where $N_xN_y$ is the number of pixels in the output image. With $N_x = N_y = N_s = 512$ and a required frame rate between ten and twenty, we find that the required number of FLOPS for real-time volume rendering of 3D cardiac ultrasound is in the order of several GFLOPS. When calculation of gradients for shading and more advanced processing is added this number continues to grow with a constant factor.

\subsection{Adaptive volume rendering}

Visibility driven visualization.

>>>>

The iterative composition scheme for front-to-back ray casting of one ray, from $s_0$ to $s_{N_s-1}$ where $s$ is the sample position and $N_s$ is the number of samples along the ray, is \cite{rtvg2006}
\begin{align}
I(s_0, s_{i+1}) &= I(s_0, s_{i}) + (1 - \alpha_{\text{acc},i}) c_{i+1}\\
\alpha_{\text{acc},i+1} &= \alpha_{\text{acc},i} + (1 - \alpha_{\text{acc},i}) \alpha(c_i+1).
\label{eq:fronToBack}
\end{align}
Here, $\alpha(v)$ is the opacity transfer function, $c_{k}$ is the color intensity of sample $s_{k}$, $\alpha_{\text{acc}}$ is the accumulated opacity, and $I(s_0, s_{N_s-1})$ is the intensity value of all the projected samples combined. If $\alpha(v)$ is close to one for one or two consecutive samples, $\alpha_{\text{acc}}$ will quickly approach a value of one. An $\alpha_{acc,k}$-value close to one means that the contribution from $c_{k+1}, \cdots, c_{N_s-1}$ to the final value, $I(s_0, s_{N_s-1})$, will be minimal. Hence, the iteration can stop at $i=k$ without loosing visible samples. This optimization technique is known as \textit{early-ray-termination}, and will significantly speed up our rendering algorithm if an opaque material is rendered.

@@

Volume rendering of a single volume has a complexity of $O(N_xN_yN_s)$, where $N_xN_y$ is the number of pixels in the output image. With $N_x = N_y = N_s = 512$ and a required frame rate for 3D cardiac ultrasound between ten and twenty, we find that the required number of FLOPS for real-time volume rendering of 3D cardiac ultrasound is in the order of several GFLOPS. When calculation of gradients for shading and more advanced processing is added per sample this number continues to grow with a constant factor.

\subsection{Adaptive volume rendering}
As for adaptive beamforming, adaptive volume rendering means to derive and use information about the volume being visualized from the volume it self. In the literature on volume rendering this is known as feature enhancement or visibility driven visualization \cite{viola2005, correa2010visibility, marchesin2010}. The common feature of these techniques is how the sample opacity is modulated in order to achieve visibility of relevant structures. If we somehow are able to classify samples as cardiac tissue we can use this information as input to an adaptive volume rendering algorithm, and by that achieve improved visibility of cardiac tissue.

In \textbf{Paper IV} we extend the work by Honigmann \textit{et al.} \cite{Honigmann2003} on a global adaptive OTF by formulating an regional adaptive OTF for visualization of cardiac ultrasound volumes. This regional adaptive OTF is based on estimated statistics of the clutter noise inside the left ventricle (the blood pool). The added complexity of this technique is the statistical estimation which is basically a second volume rendering (mean and variance projection) that has to be performed before the rendered image is produced.
